---
title: "Us police shooting dataset"
author: Rowan tarek 
output: html_document
---



### data description

The us-police-shootings data set is collected to take some insights and analyze the story about racism in America by kaggle was taken from this URL **https://www.kaggle.com/ahsen1330/us-police-shootings**
the data set includes 15 variables :

**id**: integer variable  showing Identity number.

**name** : character variable identify the name.

**date** : character variable containing the day of event.

**manner_of_death** :character variable show How they were killed.

**armed**:variables shows Weapons.

**age**:numerical variable shows age of shot person.

**gender**: ordinal  character variable showing the gender of shot person.

**race** : character variable identify race of shot person.

**city** :character variable identify Location (City) of event.

**state** :character variable identify Location (State) of event.

**signs_of_mental_illness** : character variable indicate whether there were signs of mental illness. 

**threat_level** :character variable identify the level of threats.

**flee** : character variable identify whither the person escape .

**body_camera** : character variable indicate whether policeman had body camera

**arms_category** :character variable to indicateif the person was armed



### Read/Import Data

-   read.csv() function is used to read this data and data was named as mydataset.

-   head()function is used to show the head output of data 

```{r}
mydataset<-read.csv("us-police-shootings.csv")
head(mydataset)

```

### inspect and Understand

-   variables were inspected using (ls)function.

-   dimensions of the data frame was checked using dim() function.

-   the types of of variables were summarized using summary()function

-   Data types of the variables were checked and they are in the correct format and also the levels of factor variables were checked using str()function .

-   column names were checked and column [13] was renamed from "flee" to "escape" using names()function.

-   the final structure and the data type of the data was given using str()function.


```{r}
ls(mydataset) 
dim(mydataset)
summary(mydataset)
str(mydataset)
names(mydataset)
names(mydataset)[13] <- "escape"
str(mydataset)

```

### Subsetting

-   the first 10 observations (include all variables) in data set were selected using subset()function 

-   then this subset was converted to a matrix using as.matrix() function.

-   the structure of that matrix was checked using str()function

```{r}
subset<-mydataset[1:10,]
subset
the_matrix<- as.matrix(subset)
the_matrix
str(the_matrix)

```


### Create a New Data frame

-   A new data frame was created from scratch with the given specifications (one integer variable and one ordinal variable and 10 observations) using data.frame() function .

-   the variable definition are given below :
**Patient_ID**: String variable identify the ID of patients in Doctor clinic
**Extent_of_satisfaction** :ordinal  character variable showing the satisfaction level of patient toward the clinic and the doctor 

-   new variable the_shift is created using cbind() function then this vector added to the data frame 


-   the attributes of the resulting data frame was checked using attributes() function .

-   dimensions of the data set is checked using dim()

```{r}
Patient_ID<-c(1,12,105,3,108,104,407,156,189,111)
Extent_of_satisfaction <-c("Indifferent","dissatisfied","Satisfied","Very satisfied ","Satisfied","Indifferent","Dissatisfied ","Satisfied","Very dissatisfied","Satisfied")

df<-data.frame(Patient_ID,Extent_of_satisfaction)
df
class(df)
factor_satisfaction<-factor(df$Extent_of_satisfaction)
Extent_of_satisfaction <- factor(Extent_of_satisfaction,ordered = TRUE,levels =c("Very satisfied","Satisfied","Indifferent","Dissatisfied","Very dissatisfied"))
Extent_of_satisfaction
str(df)

the_shift <-c("night","day","night","night","day","night","day","night","night","day")
new_df<-df
new_df <- cbind(df, the_shift = the_shift)
new_df
attributes(new_df)
dim(new_df)

                                     
```